---
title: "Median and Mean CPUE"
author: "Alexander Steer-Milnes"
date: "28/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(lubridate)
library(rmarkdown)
CLandings_data <- read.csv("data/sample_landing_data_clean.csv")
```

## Comparing Median and Mean CPUE

First we need to calculate the catch per unit efficiency. CPUE is mesured in kg/h so we need to change our date somewhat before calculating this.

we'll start with the previously cleaned ladings data and work from there.

```{r CLD, echo=FALSE}
summary(CLandings_data)
```

Firstly our data has weight in grams, so we want to change this to Kilograms.
Secondly we can group by year and trip ID so we can calculate the CPUE for each year.
Thirdly we can actually calculate CPUE by taking the total weight of the catch and dividing by the number of hours for each trip.

```{r g to kg, echo=FALSE}
CPUE_data <- CLandings_data %>%
  mutate(Weight_kg = Weight_g / 1000) %>%
  group_by(Year,Trip_ID) %>%
  summarize(Trip_CPUE = sum(Weight_kg) / mean(Effort_Hours))
CPUE_data
```

Now we want to compare both the median and mean of the CPUE for each year.

```{r MedianMean, echo=FALSE}
CPUE_Median <- CPUE_data %>%
  group_by(Year) %>%
  summarize(Median_CPUE = median(Trip_CPUE))
CPUE_Mean <- CPUE_data %>%
  group_by(Year) %>%
  summarize(Mean_CPUE = mean(Trip_CPUE))
CPUE_Median
CPUE_Mean
```

This is easier to visulaise in a plot.

```{r MedianvsMean, echo=FALSE}
ggplot() +
  geom_point(data = CPUE_Median, aes(x = Year, y = Median_CPUE), color = "blue") +
  geom_line(data = CPUE_Median, aes(x = Year, y = Median_CPUE), color = "blue") +
  geom_point(data = CPUE_Mean, aes(x = Year, y = Mean_CPUE), color = "red") +
  geom_line(data = CPUE_Mean, aes(x = Year, y = Mean_CPUE), color = "red") +
  ylab("CPUE") +
  ggtitle("Median CPUE vs Mean CPUE") +
  scale_x_continuous(breaks = seq(2003, 2014, by = 1))
```

We can see that between 2005 and 2012 the Mean (in red) and median (in blue) are very close. However in 2003, 2004, 2013 and 2014 the mean is significantly higher than the median this might suggest that some trips were significantly higher in CPUE that the rest for those years
